Proposal: A Natural Language Programming Toolkit for Python

### 2001-03-03 technical/proposal/proposal.tex
Spell-checked.
svn/trunk@16

CIS-530 is an introductory course in natural language processing
(NLP), which is intended to be accessable to both linguists and
computer scientists \cite{cis530}. In such a class, it is instructive
for students to build real NLP systems.  By building real systems,
students can gain an in-depth understanding of the issues involved in
particular aspects of NLP.

However, the students do not necessarily have any experience
programming or designing software systems.  Thus, when students
attempt to build NLP systems, they may spend a signifigant amount of
their time on programming tasks that are not directly related to the
course of the content.

@@

  \item It will provide the students with an infrastructure for
  building natural language systems, by providing consistant and
  well-defined interfaces for different components.  Thus, students
  will not have to spend their time thinking about system design
  issues.
\end{itemize}

@@

\item[Consistancy] The toolkit should use consistant data structures
and interfaces.

\item[Extensibility] The toolkit should easily accomodate new
components, whether those components replicate or extend the toolkit's
existing functionality.  Thus, the toolkit's design should be modular,
with simple and well-defined interfaces between modules.  The toolkit
should also be structured in such a way that it is obvious where new
extensions would fit into the toolkit's infrastructure.

\item[Documentation] The toolkit, its data structures, and its
implementation all need to be carefully and thouroughly documented.

@@

\item[Modularity] The interaction between different components of the
toolkit should be kept to a minimum.  In particular, it should be
possible to use complete individaul projects using small parts of the
toolkit, without worrying about how they interact with the rest of the
toolkit.  This will allow students to learn to use the toolkit
incrementally throughout the course.  Modularity also makes it easier
to change or extend the toolkit.
\end{description}

@@

The natural langauge processing toolkit will be implemented by a
collection of independant modules\footnote{These modules do not
necessarily correspond to Python \texttt{module}s.}.  These modules are
organized according to two primary axes:

@@

\subsubsection{Dependancies}

In general, data-oriented modules should be self-contained, and should
exhibit very little inter-dependancy.  Data-oriented modules never
depend on task-oriented modules.

Task-oriented modules should also be relatively self-contained, with
very little inter-dependancy.  However, each task-oriented modules
will usually depend on a variety of data-oriented modules.  When
possible, the number of data-oriented modules that a task-oriented
module depends on should be kept low.  This allows students to use
task-oriented modules without needing to learn how to use a large
number of other modules.

@@

\vspace{1.2mm}\noindent
\begin{tabular}{||p{.21\textwidth}p{.7\textwidth}}
  \textbf{Token} & An Occurance of a \texttt{TokenType}
  within a text.  A \texttt{token} consists of a \texttt{TokenType}
  and a \texttt{TextLocation}. \\
\end{tabular}
\vspace{1mm}

@@

When learning to use the toolkit, students intuitions about each
entity in the toolkit will be guided by the entities names'.  It is
therefore important to use consistant, correct, and intuitive
nomenclature when naming the entities defined by the toolkit.
The documentation should be careful to always use correct
nomenclature; in particular, it should never conflate two concepts
with the same name.

@@

Type checking is an imporant safety device, that allows programmers to
quickly detect the locations of errors in their code.  This is
especially important for novice programmers, who have less experience
with debugging.

@@

  \item[Mapping and Filtering] Mapping and filtering provide a
  consise way of expressing transformations on lists.  However, they
  are fairly difficult to learn, and generally involve the use of
  \texttt{lambda} functions, so the toolkit will not use them.

@@

Technical reports will be available as latex files, dvi files, and
postscript files.  Tutorials and reference documentation will be
available as HTML files, accessble over the web.

@@

Throughout the course of developing the toolkit, I will write a number
of technical reports, which are intended to explain and justify
specific aspects of the toolkit's design or implementation.  This
document is an example of a technical report.  Students should not
need to read the technical reports to use the toolkit.  However,
students can consult these reports if they would like further
information about how the tookit is designed and why it is designed
that way.

@@

Reference documentation gives definitions for each module, interface,
class, method, function, and variable defined by the toolkit.  These
definitions should be both precise and consise.  I currently plan to
produce the reference documentation using Epydoc \cite{epydoc}.
Epydoc is a tool for automatically converting python documentation
strings to HTML pages.  Epydoc will generate the following HTML pages:

\begin{itemize}
  \item A page for each python \texttt{module}.  This page includes an
  overview of the \texttt{module}; a list of the \texttt{classs}es
  defined by the \texttt{module}; a list of the \texttt{function}s
  defined by the \texttt{module}, and a detailed description of each
  \texttt{function}; a list of the \texttt{variable}s defined by the
  \texttt{module}, and a detailed description of each
  \texttt{variable}.

  \item A page for each python \texttt{class}.  This page includes an
  overview of the \texttt{classs}; a list of the \texttt{method}s
  defined by the \texttt{classs}, and a detailed description of each
  \texttt{method}; a list of the instance \texttt{variable}s defined
  by the \texttt{classs}, and a detailed description of each instance
  \texttt{variable}; a list of the \texttt{classs} \texttt{variable}s
  defined by the \texttt{classs}, and a detailed description of each
  \texttt{classs} \texttt{variable}.

  \item A page containing the toolit's complete \texttt{classs}
  hierarchy.

  \item A page containing an index of every \texttt{module},
  \texttt{classs}, \texttt{method}, \texttt{function}, and
  \texttt{variable} defined by the toolkit.

@@

As part of the preliminary design process, I desinged and implemented
six prototype modules.  These modules are summarized in figures
\ref{fig:mod-data} and \ref{fig:mod-task}, and described in the
following sections.  Most of these descriptions are copied directly from
the modules' reference documentation.  For more information, see the
reference documentation 
at ``\texttt{http://www.cis.upenn.edu/$\sim$edloper/nltk/nltk.html}''.

\begin{figure}
\noindent
\begin{centering}
\begin{boxedminipage}{.95\textwidth}
\begin{itemize}
  \item \textbf{Sets}: Encodes the mathmatical notion of a ``finite set.''
  \begin{itemize}
    \item \textit{Set}: A finite set.
  \end{itemize}

@@

  \item \textbf{Probability}: Encodes data structures associated with
        the mathmatical notion of probability, such as events and
        frequency distributions.
  \begin{itemize}
    \item Sample: Encodes the mathmatical notion of a
          ``sample.''  This is actually not implemented as a class or
          an interface -- (almost) anything can be a Sample.
    \item \textit{EventI}:
         A (possibly infinite) set of samples.
    \item \textit{FreqDistI}:
          The frequency distribution of a collection of samples.
    \item \textit{ProbDistI}:
          A probability distribution, typically derived from a
          frequency distribution (e.g., using ELE).
  \end{itemize}
\end{itemize}
\end{boxedminipage}\\
\end{centering}
 \caption[Data-oriented module prototypes]{Data-oriented module
 prototypes.  Under each module, the interfaces and top-level classes
 defined by that module are listed.}
\label{fig:mod-data}
\end{figure}

@@

    \begin{itemize}
      \item For all elements $e_i$, $e_j$ added to the \texttt{Set},
           $e_i=e_j$ if and only if $e_j=e_i$.  This should always be the
           case as long as the elements in the \texttt{Set} use
           well-defined comparison functions.  An example where it
           would not be the case would be if ei defined
           \texttt{\_\_cmp\_\_()} to always return 0, and ej
           defined \texttt{\_\_cmp\_\_()} to always return $-1$.

@@

    A \texttt{TokenTypeI} is single unit of text, such as a word or a
    punctuation mark.  The precise defintion of what counts as a token
    type will vary for different analyses.  For example, ``bank'' and
    ``run'' might be tokens for one analysis, while ``bank/N'' and
    ``bank/V'' might be tokens for another analysis.

@@

    A \texttt{Token} is an occurance of a single unit of text, such as
    a word or a punctuation mark.  A \texttt{Token} consists of a token type
    and a source.  The token type is the unit of text (e.g., a
    specific word).  The source is the position at which this token
    occured in the text.

    The precise defintion of what counts as a token type, or unit of
    text, will vary for different analyses.  For example, ``bank'' and 
    ``run'' might be tokens for one analysis, while ``bank/N'' and
    ``bank/V'' might be tokens for another analysis.

@@

The probability module provides data structures associated with the
mathmatical notion of probability.  The following mathmatical terms
are useful for thinking about probability:

@@

The toolkit currently defines three interfaces for handling the
mathmatical notion of probability: the \texttt{EventI} interface; the
\texttt{FreqDistI} interface; and the \texttt{ProbDistI} interface.

@@

    A \texttt{FreqDist} represents the frequency distribution for the
    outcomes of an experiment.  A frequency distribution records the
    number of times each outcome of an experiment has occured.  For
    example, a frequency distribution could be used to record the
    frequency of each token in a document.  Formally, a frequency
    distribution can be defined as a function mapping from samples to
    the number of times that sample occured as an outcome.

@@

    Frequency distributions are required to implement the methods
    \texttt{inc()}, \texttt{N()}, \texttt{count()},
    \texttt{freq()}, \texttt{max()}, \texttt{cond\_freq()}, 
    and \texttt{cond\_max()}.  In the future, this list may be
    exapanded, and optional methods may be added.

@@

    Probability distributions are required to implement the methods
    \texttt{prob()} and \texttt{cond\_prob()},
    \texttt{max()}, and \texttt{cond\_max()}.  In the future,
    this list may be exapanded, and optional methods may be added.

@@

The process of assigning a descriptive tag to each \texttt{Token} in a
list of \texttt{Token}s is called \emph{tagging}.  Often, tagging is
used to assign a part-of-speech tag to each \texttt{Token} in a list
of \texttt{Token}s.  Particular \texttt{Taggers} may attatch different
types of tags to the \texttt{Token}s in the list, or may use different
algorithms for deciding which tag to attatch.

@@

%################################################################
%#  BiIBLIOGRAPHY
%################################################################
\begin{thebibliography}{1}

>>>>

CIS-530 is an introductory course in natural language processing
(NLP), which is intended to be accessible to both linguists and
computer scientists \cite{cis530}. In such a class, it is instructive
for students to build real NLP systems.  By building real systems,
students can gain an in-depth understanding of the issues involved in
particular aspects of NLP.

However, the students do not necessarily have any experience
programming or designing software systems.  Thus, when students
attempt to build NLP systems, they may spend a significant amount of
their time on programming tasks that are not directly related to the
course of the content.

@@

  \item It will provide the students with an infrastructure for
  building natural language systems, by providing consistent and
  well-defined interfaces for different components.  Thus, students
  will not have to spend their time thinking about system design
  issues.
\end{itemize}

@@

\item[Consistency] The toolkit should use consistent data structures
and interfaces.

\item[Extensibility] The toolkit should easily accommodate new
components, whether those components replicate or extend the toolkit's
existing functionality.  Thus, the toolkit's design should be modular,
with simple and well-defined interfaces between modules.  The toolkit
should also be structured in such a way that it is obvious where new
extensions would fit into the toolkit's infrastructure.

\item[Documentation] The toolkit, its data structures, and its
implementation all need to be carefully and thoroughly documented.

@@

\item[Modularity] The interaction between different components of the
toolkit should be kept to a minimum.  In particular, it should be
possible to use complete individual projects using small parts of the
toolkit, without worrying about how they interact with the rest of the
toolkit.  This will allow students to learn to use the toolkit
incrementally throughout the course.  Modularity also makes it easier
to change or extend the toolkit.
\end{description}

@@

The natural language processing toolkit will be implemented by a
collection of independent modules\footnote{These modules do not
necessarily correspond to Python \texttt{module}s.}.  These modules are
organized according to two primary axes:

@@

\subsubsection{Dependencies}

In general, data-oriented modules should be self-contained, and should
exhibit very little inter-dependency.  Data-oriented modules never
depend on task-oriented modules.

Task-oriented modules should also be relatively self-contained, with
very little inter-dependency.  However, each task-oriented modules
will usually depend on a variety of data-oriented modules.  When
possible, the number of data-oriented modules that a task-oriented
module depends on should be kept low.  This allows students to use
task-oriented modules without needing to learn how to use a large
number of other modules.

@@

\vspace{1.2mm}\noindent
\begin{tabular}{||p{.21\textwidth}p{.7\textwidth}}
  \textbf{Token} & An occurance of a \texttt{TokenType}
  within a text.  A \texttt{token} consists of a \texttt{TokenType}
  and a \texttt{TextLocation}. \\
\end{tabular}
\vspace{1mm}

@@

When learning to use the toolkit, students intuitions about each
entity in the toolkit will be guided by the entities names'.  It is
therefore important to use consistent, correct, and intuitive
nomenclature when naming the entities defined by the toolkit.
The documentation should be careful to always use correct
nomenclature; in particular, it should never conflate two concepts
with the same name.

@@

Type checking is an important safety device, that allows programmers to
quickly detect the locations of errors in their code.  This is
especially important for novice programmers, who have less experience
with debugging.

@@

  \item[Mapping and Filtering] Mapping and filtering provide a
  concise way of expressing transformations on lists.  However, they
  are fairly difficult to learn, and generally involve the use of
  \texttt{lambda} functions, so the toolkit will not use them.

@@

Technical reports will be available as latex files, dvi files, and
postscript files.  Tutorials and reference documentation will be
available as HTML files, accessible over the web.

@@

Throughout the course of developing the toolkit, I will write a number
of technical reports, which are intended to explain and justify
specific aspects of the toolkit's design or implementation.  This
document is an example of a technical report.  Students should not
need to read the technical reports to use the toolkit.  However,
students can consult these reports if they would like further
information about how the toolkit is designed and why it is designed
that way.

@@

Reference documentation gives definitions for each module, interface,
class, method, function, and variable defined by the toolkit.  These
definitions should be both precise and concise.  I currently plan to
produce the reference documentation using Epydoc \cite{epydoc}.
Epydoc is a tool for automatically converting python documentation
strings to HTML pages.  Epydoc will generate the following HTML pages:

\begin{itemize}
  \item A page for each python \texttt{module}.  This page includes an
  overview of the \texttt{module}; a list of the \texttt{class}es
  defined by the \texttt{module}; a list of the \texttt{function}s
  defined by the \texttt{module}, and a detailed description of each
  \texttt{function}; a list of the \texttt{variable}s defined by the
  \texttt{module}, and a detailed description of each
  \texttt{variable}.

  \item A page for each python \texttt{class}.  This page includes an
  overview of the \texttt{class}; a list of the \texttt{method}s
  defined by the \texttt{class}, and a detailed description of each
  \texttt{method}; a list of the instance \texttt{variable}s defined
  by the \texttt{class}, and a detailed description of each instance
  \texttt{variable}; a list of the \texttt{class} \texttt{variable}s
  defined by the \texttt{class}, and a detailed description of each
  \texttt{class} \texttt{variable}.

  \item A page containing the toolkit's complete \texttt{class}
  hierarchy.

  \item A page containing an index of every \texttt{module},
  \texttt{class}, \texttt{method}, \texttt{function}, and
  \texttt{variable} defined by the toolkit.

@@

As part of the preliminary design process, I designed and implemented
six prototype modules.  These modules are summarized in figures
\ref{fig:mod-data} and \ref{fig:mod-task}, and described in the
following sections.  Most of these descriptions are copied directly from
the modules' reference documentation.  For more information, see the
reference documentation 
at ``\texttt{http://www.cis.upenn.edu/$\sim$edloper/nltk/nltk.html}''.

\begin{figure}
\noindent
\begin{centering}
\begin{boxedminipage}{.95\textwidth}
\begin{itemize}
  \item \textbf{Sets}: Encodes the mathematical notion of a ``finite set.''
  \begin{itemize}
    \item \textit{Set}: A finite set.
  \end{itemize}

@@

  \item \textbf{Probability}: Encodes data structures associated with
        the mathematical notion of probability, such as events and
        frequency distributions.
  \begin{itemize}
    \item Sample: Encodes the mathematical notion of a
          ``sample.''  This is actually not implemented as a class or
          an interface -- (almost) anything can be a Sample.
    \item \textit{EventI}:
         A (possibly infinite) set of samples.
    \item \textit{FreqDistI}:
          The frequency distribution of a collection of samples.
    \item \textit{ProbDistI}:
          A probability distribution, typically derived from a
          frequency distribution (e.g., using ELE).
  \end{itemize}
\end{itemize}
\end{boxedminipage}\\
\end{centering}
 \caption[Data-oriented module prototypes]{Data-oriented module
 prototypes.  Under each module, the interfaces and top-level classes
 defined by that module are listed.}
\label{fig:mod-data}
\end{figure}

@@

    \begin{itemize}
      \item For all elements $e_i$, $e_j$ added to the \texttt{Set},
           $e_i=e_j$ if and only if $e_j=e_i$.  This should always be the
           case as long as the elements in the \texttt{Set} use
           well-defined comparison functions.  An example where it
           would not be the case would be if $e_i$ defined
           \texttt{\_\_cmp\_\_()} to always return 0, and $e_j$
           defined \texttt{\_\_cmp\_\_()} to always return $-1$.

@@

    A \texttt{TokenTypeI} is single unit of text, such as a word or a
    punctuation mark.  The precise definition of what counts as a token
    type will vary for different analyses.  For example, ``bank'' and
    ``run'' might be tokens for one analysis, while ``bank/N'' and
    ``bank/V'' might be tokens for another analysis.

@@

    A \texttt{Token} is an occurance of a single unit of text, such as
    a word or a punctuation mark.  A \texttt{Token} consists of a token type
    and a source.  The token type is the unit of text (e.g., a
    specific word).  The source is the position at which this token
    occurred in the text.

    The precise definition of what counts as a token type, or unit of
    text, will vary for different analyses.  For example, ``bank'' and 
    ``run'' might be tokens for one analysis, while ``bank/N'' and
    ``bank/V'' might be tokens for another analysis.

@@

The probability module provides data structures associated with the
mathematical notion of probability.  The following mathematical terms
are useful for thinking about probability:

@@

The toolkit currently defines three interfaces for handling the
mathematical notion of probability: the \texttt{EventI} interface; the
\texttt{FreqDistI} interface; and the \texttt{ProbDistI} interface.

@@

    A \texttt{FreqDist} represents the frequency distribution for the
    outcomes of an experiment.  A frequency distribution records the
    number of times each outcome of an experiment has occurred.  For
    example, a frequency distribution could be used to record the
    frequency of each token in a document.  Formally, a frequency
    distribution can be defined as a function mapping from samples to
    the number of times that sample occurred as an outcome.

@@

    Frequency distributions are required to implement the methods
    \texttt{inc()}, \texttt{N()}, \texttt{count()},
    \texttt{freq()}, \texttt{max()}, \texttt{cond\_freq()}, 
    and \texttt{cond\_max()}.  In the future, this list may be
    expanded, and optional methods may be added.

@@

    Probability distributions are required to implement the methods
    \texttt{prob()} and \texttt{cond\_prob()},
    \texttt{max()}, and \texttt{cond\_max()}.  In the future,
    this list may be expanded, and optional methods may be added.

@@

The process of assigning a descriptive tag to each \texttt{Token} in a
list of \texttt{Token}s is called \emph{tagging}.  Often, tagging is
used to assign a part-of-speech tag to each \texttt{Token} in a list
of \texttt{Token}s.  Particular \texttt{Taggers} may attach different
types of tags to the \texttt{Token}s in the list, or may use different
algorithms for deciding which tag to attach.

@@

%################################################################
%#  BIBLIOGRAPHY
%################################################################
\begin{thebibliography}{1}
